### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@0bfcf11d2028cedf01a37ba47a7b6385fe6e60cd 🚀

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs.\textsuperscript{1} Although host richness is often used as a
superficial proxy for spillover risk,\textsuperscript{2,3} these
approaches oversimplify the relevant interspecific heterogeneity in
immunology, behavior, and other traits, and therefore overlook unique
host pools that allow for the rapid evolution of highly divergent
viruses.\textsuperscript{4} In the case of generalist pathogens like
betacoronaviruses, there is conceptual and empirical support to the idea
that these community-level mechanisms are even more
important,\textsuperscript{5} particularly given that cross-species
transmission may, as a rule, structure viral evolution more than
co-divergence with hosts.\textsuperscript{6} This creates a disconnect
between coevolutionary theory (including empirical evidence from
virology) and most existing ecological frameworks for mapping spillover
risk.

@@

Most previous work has assumed that coronavirus biogeography is driven
by coevolutionary regimes that form at finer taxonomic scales, treating
the presence or richness of key bat host groups as predictive of these
viruses' distribution.\textsuperscript{2,3} Projecting bat and
betacoronavirus phylogeny over space (fig.~\ref{fig:biogeo}), we find
further support for the idea that bat community assembly is directly
responsible for spatial heterogeneity in viral coevolutionary regimes.
The distinct groupings (represented by different colors, symbolizing
positions in a subspace formed by the first two phylogenetic principal
components) are essentially equivalent between the two groups, and can
be coarsely delineated as (1) south and southeast Asia, (2) east Asia
(including northern China), west Asia, and the Mediterranean coast; (3)
Eurasia above a northing of 40; and (4) Africa and south America. In
some cases, this diverges from expectations about coronavirus
biogeography: for example, previous work has rarely flagged India as a
region of concern, but for both bats and betacoronaviruses, the
subcontinent falls into the same phylogeographic regions as the
southeast Asian peninsula (and indeed, the region is home to known bat
hosts of nobecoviruses, sarbecoviruses, and
merbecoviruses).\textsuperscript{3}

@@

Overall, these results suggest that, although there are unique hotspots
of viral adaptive radiation, biogeographic regions are mostly consistent
between bats and their viruses. This may be surprising, given that
cospeciation plays a minor role in coronavirus
diversification,\textsuperscript{2} which could theoretically allow for
broad divergence in their biogeography. However, host jumps at the
family level or higher are relatively rare and significant events in
coronavirus evolutionary history;\textsuperscript{2,17} as a result, the
mosaic of betacoronavirus phylogeography is assembled from a set of
overlapping smaller coevolutionary systems, superimposed in space and
filtered by the importance of different subgroups in local host
communities. For example, the most speciose and cosmopolitan family of
bats, the vesper bats (Vespertilionidae), are considered the primary
reservoir of merbecoviruses;\textsuperscript{3,17} but in the Americas,
where merbecoviruses are the only lineage present, they have only been
found in other bat taxa. At the coarsest scale, these heterogeneities
are lost, and betacoronavirus biogeography tracks the deep rifts in bat
evolutionary history---but within these broad regions, the component
coevolutionary systems can have very different dynamics.

>>>>

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs.\textsuperscript{1} Although host richness is often used as a
superficial proxy for spillover risk,\textsuperscript{2,3} these
approaches oversimplify the relevant interspecific heterogeneity in
immunology, behavior, and other traits, and therefore overlook unique
host pools that allow for the rapid evolution of highly divergent
viruses.\textsuperscript{4} In the case of generalist pathogens like
betacoronaviruses, there is conceptual and empirical support to the idea
that these community-level mechanisms are even more
important,\textsuperscript{5} particularly given that cross-species
transmission may, as a rule, structure viral evolution more than
co-divergence with hosts.\textsuperscript{6} This creates a disconnect
between coevolutionary theory and most existing ecological frameworks
for mapping spillover risk.

@@

Most previous work has assumed that coronavirus biogeography is driven
by coevolutionary regimes that form at finer taxonomic scales, treating
the presence or richness of key bat host groups as predictive of these
viruses' distribution.\textsuperscript{2,3} Projecting bat and
betacoronavirus phylogeny over space (fig.~\ref{fig:biogeo}), we find
further support for the idea that bat community assembly is directly
responsible for spatial heterogeneity in viral coevolutionary regimes.
The distinct groupings (represented by different colors, symbolizing
positions in a subspace formed by the first two phylogenetic principal
components) are essentially equivalent between the two groups, and can
be coarsely delineated as (1) south and southeast Asia, (2) east Asia
(including northern China), west Asia, and the Mediterranean coast; (3)
Eurasia above a northing of 40; and (4) Africa and south America. In
some cases, this diverges from expectations about coronavirus
biogeography: for example, previous work has rarely flagged India as a
region of interest, but for both bats and betacoronaviruses, the
subcontinent falls into the same regions as the southeast Asian
peninsula (and indeed, the region is home to known bat hosts of
nobecoviruses, sarbecoviruses, and merbecoviruses).\textsuperscript{3}

@@

Overall, these results suggest that, although there are unique hotspots
of viral adaptive radiation, biogeographic regions are mostly consistent
between bats and their viruses. This may be surprising, given that
cospeciation plays a minor role in coronavirus
diversification,\textsuperscript{2} which could theoretically allow for
broad divergence in their biogeography. However, host jumps at the
family level or higher are relatively rare and significant events in
coronavirus evolutionary history;\textsuperscript{2,17} as a result, the
mosaic of betacoronavirus phylogeography is assembled from a set of
overlapping smaller coevolutionary systems, superimposed in space and
filtered by the importance of different subgroups in local host
communities. For example, the most speciose and cosmopolitan family of
bats, the vesper bats (Vespertilionidae), are considered the primary
reservoir of merbecoviruses;\textsuperscript{3,17} but in the Americas,
where merbecoviruses are the only lineage present, they have only been
found in other bat taxa. At the coarsest scale, these heterogeneities
are lost, and betacoronavirus biogeography tracks the deep rifts in bat
evolutionary history---but within broad regions, the component
coevolutionary systems may have very different dynamics.
